Dynamic Walking and Whole-Body Motion Planning for Humanoid Robots: an Integrated Approach

### 2012-06-20 paper.tex
Algo 2 returns a configuration and Algo 3 uses q'' instead of q'.
Add a column in table to make result look less miserable.

The problem of inverse kinematics for a humanoid robot, or any articulated
structure, is to compute a joint motion to achieve an end-effector pose. As the
robots we deal with are redundant, it is natural to take advantage of
this redundancy by specifying multiple tasks, potentially with
different priorities. This problem has been widely studied in robotics
planning and control literature, and many jacobian-based solutions have been
proposed, among which 
\cite{nakamura1986iks}, \cite{siciliano1991gfm},
\cite{baerlocher1998tpf} and \cite{khatib2004wbd}.
Obstacle avoidance can be taken into account with similar methods. To
do so, one has to include the obstacles as  constraints to
satisfy, see for example \cite{kanehiro2008lca}.
These methods are prone to fall into local minima, global motion 
planning is needed to overcome this limitation.
\cite{TouGieGoe2007} Propose a motion generation method where tasks follow
trajectories define by cubic B-splines. The whole-body motion is optimized
with respect to the control point positions. This method can take into account
collision with simple obstacles.

@@

\begin{algorithm}
\caption{constraint-solver($q$, f, $\varepsilon$): find $q$ such that $f(q) = 0$}
\label{algo:newton}
\begin{algorithmic}
\STATE $\alpha=.1$, $i=0$, $\alpha_{max} = .95$
\WHILE{$\|f(q)\| > \varepsilon$ and $i\leq 20$}
\STATE {\color{red} // $\left(.\right)^{+}$ denotes the Moore-Penrose pseudo-inverse}
\STATE $q \leftarrow$ $q - \alpha \left(\frac{\partial f}{\partial q}(q)\right)^{+} f(q)$
\STATE $i$ $\leftarrow$ $i+1$
\STATE {\color{red}// Make $\alpha$ tend toward $\alpha_{max}$}
\STATE $\alpha \rightarrow \alpha_{max} - .8*(\alpha_{max} - \alpha)$
\ENDWHILE
\IF {$\|f(q)\| \leq \varepsilon$}
\STATE return success
\ELSE
\STATE return failure
\ENDIF
\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}[h]
  \caption{Constrained-Extend($\mathcal{T},q_{near},q_{rand},f,\epsilon$)}
  \label{alg:constrained}
  \begin{algorithmic}
    \STATE {\color{red} // $I$ denotes the identity matrix.}
    \STATE $q_{rand}' \leftarrow q_{near} +   
    \left(I -  
    \left(\frac{\partial f}{\partial q}(q_{near})\right)^{+}
    \frac{\partial f}{\partial q}(q_{near})\right) (q_{rand} - q_{near})$
    \IF {$\text{constraint-solver}(q_{rand}',f,\epsilon) = success$} 
    \STATE RRT::Extend($\mathcal{T},q_{near},q_{rand}'$)
    \ENDIF
  \end{algorithmic}
\end{algorithm}

@@

We have run this set of motion planning problems 20 times. The results are compiled in Table~\ref{table:reaching}. We can notice that optimization dominates the total time of computation. This suggests that improvement can be easily achieved by using numerical optimization techniques. This is on-going work.
\begin{table}
\begin{tabular}{lrrr}
& min & max & average \\
number of nodes & 43.00 & 481.00 & 102.70 \\
goal generation time (seconds) \hspace {1cm} & 1.00 & 1.56 & 1.22 \\
planning time (seconds)& 67.36 & 376.84 & 134.28 \\
optimization time (seconds) & 204.22 & 671.26 & 521.47 \\
total times (seconds) & 468.23 & 864.68 & 656.97
\end{tabular}
\caption {Experimental results on 20 runs: 3 motion planning problems and 2 goal generations.}
\label{table:reaching}
\end{table}

>>>>

The problem of inverse kinematics for a humanoid robot, or any articulated
structure, is to compute a joint motion to achieve an end-effector pose. As the
robots we deal with are redundant, it is natural to take advantage of
this redundancy by specifying multiple tasks, potentially with
different priorities. This problem has been widely studied in robotics
planning and control literature, and many jacobian-based solutions have been
proposed, among which 
\cite{nakamura1986iks}, \cite{siciliano1991gfm},
\cite{baerlocher1998tpf} and \cite{khatib2004wbd}.
Obstacle avoidance can be taken into account with similar methods. To
do so, one has to include the obstacles as  constraints to
satisfy, see for example \cite{kanehiro2008lca}.
These methods are prone to fall into local minima, global motion 
planning is needed to overcome this limitation.
\cite{TouGieGoe2007} propose a motion generation method where tasks follow
trajectories define by cubic B-splines. The whole-body motion is optimized
with respect to the control point positions. This method can take into account
collision with simple obstacles.

@@

\begin{algorithm}
\caption{constraint-solver($q$, f, $\varepsilon$): find $q$ such that $f(q) = 0$}
\label{algo:newton}
\begin{algorithmic}
\STATE $\alpha=.1$, $i=0$, $\alpha_{max} = .95$
\WHILE{$\|f(q)\| > \varepsilon$ and $i\leq 20$}
\STATE {\color{red} // $\left(.\right)^{+}$ denotes the Moore-Penrose pseudo-inverse}
\STATE $q \leftarrow$ $q - \alpha \left(\frac{\partial f}{\partial q}(q)\right)^{+} f(q)$
\STATE $i$ $\leftarrow$ $i+1$
\STATE {\color{red}// Make $\alpha$ tend toward $\alpha_{max}$}
\STATE $\alpha \rightarrow \alpha_{max} - .8*(\alpha_{max} - \alpha)$
\ENDWHILE
\IF {$\|f(q)\| \leq \varepsilon$}
\STATE return $q$
\ELSE
\STATE return failure
\ENDIF
\end{algorithmic}
\end{algorithm}

@@

\begin{algorithm}[h]
  \caption{Constrained-Extend($\mathcal{T},q_{near},q_{rand},f,\epsilon$)}
  \label{alg:constrained}
  \begin{algorithmic}
    \STATE {\color{red} // $I$ denotes the identity matrix.}
    \STATE $q_{rand}' \leftarrow q_{near} +   
    \left(I -  
    \left(\frac{\partial f}{\partial q}(q_{near})\right)^{+}
    \frac{\partial f}{\partial q}(q_{near})\right) (q_{rand} - q_{near})$
    \IF {$\text{constraint-solver}(q_{rand}',f,\epsilon) = q_{rand}''$} 
    \STATE RRT::Extend($\mathcal{T},q_{near},q_{rand}''$)
    \ENDIF
  \end{algorithmic}
\end{algorithm}

@@

We have run this set of motion planning problems 20 times. The results are compiled in Table~\ref{table:reaching}. We can notice that optimization dominates the total time of computation. This suggests that improvement can be easily achieved by using numerical optimization techniques. This is on-going work.
\begin{table}
\begin{tabular}{lrrrr}
& min & max & average & average \\ 
&&&& per problem \\
number of nodes & 43.00 & 481.00 & 102.70 \\
goal generation time \hspace {1.cm} & 1.00 & 1.56 & 1.22\\
planning time& 67.36 & 376.84 & 134.28 & 44.76\\
optimization time & 204.22 & 671.26 & 521.47 & 173.82 \\
total times & 468.23 & 864.68 & 656.97 &
\end{tabular}
\caption {Experimental results on 20 runs: 3 motion planning problems and 2 goal generations. Times are expressed in seconds.}
\label{table:reaching}
\end{table}
